---
title: "A deeper dive into The Baltimore Banner poll"
author: "Nick Thieme"
date: '2022-06-16'
output: github_document
---
## What the Baltimore Now polls tells us about how city residents share opinions

### Library and data load

This is R code that goes along with The Baltimore Banner story "What the Baltimore Now poll tells us about how city residents share opinions" available at: INSERT LINK

The data analysis is relatively straightforward throughout, using the srvyr package to do the heavy lifting of accounting for polling weights in the calculations of standard errors. We use the survey package just once for calculating a correlation matrix.

```{r setup, include=FALSE}
library(haven)
library(tidyverse)
library(survey)
library(srvyr)
library(jtools)
library(weights)
library(ggcorrplot)
library(janitor)
library(dendsort)
library(seriation)
library(gplots)
library(RColorBrewer)
library(tidycensus)
library(lubridate)
library(sf)
library(vctrs)

D_poll_results = read_sav("~/Desktop/banner_projects/banner_bmore_poll/data/Baltimore_City_Poll_Baltimore Banner.sav")
D_poll_results$scott %>% table %>% {./sum(.)}

surv_design <- svydesign(id = ~1, weights = ~D_poll_results$weight, data = D_poll_results)
srs_design<-D_poll_results %>% as_survey_design(ids =1, weight = weight)
```

### Correlation plot

There's quite a bit of interesting structure in the poll responses and this structure guides a good bit of the ultimate story. By thresholding the correlation matrix and looking for blocks of questions that poll together, we can find some interesting relationships between different questions and Baltimoreans' responses.

```{r pressure, echo=FALSE}
corr_gram<-svycor(~track + opt+samosby+presmosby+scott+police+policecom+ceo+scott_fav+afford+
                    litter+ pub_transit+job_opp+crime+taxrate+grade+condition+unsafe+fund_edu+
                    standards+local_sch+ teacher+safe_city+safe_nb+misconduct+treat+protect+
                    few_dept+allocate+ harsh+prosecution+presence+gun+pol_fund+vacant+leaving, 
                  design = surv_design, digits = 4)

corr_gram[[1]][which(corr_gram[[1]]<.1)]<-0
corr_gram[[1]][which(corr_gram[[1]]>=.1&corr_gram[[1]]<.2)]<-.15
corr_gram[[1]][which(corr_gram[[1]]>=.2&corr_gram[[1]]<.3)]<-.25
corr_gram[[1]][which(corr_gram[[1]]>=.3&corr_gram[[1]]<.4)]<-.5
corr_gram[[1]][which(corr_gram[[1]]>=.4&corr_gram[[1]]<.9)]<-.75

ggcorrplot(corr_gram[[1]])
```

This shows three main blocks of questions that poll together. One related to politicians, generally, and the Mosbys specifically, another related to questions about the schools, and a third block related to criminal justice and prosecution of crimes. Looking at the topline findings shows that these blocks aren't the only things Baltimore cares about, there are other subjects that the city has even stronger feelings about. Instead, what it shows us, is that "opinion groups" have formed around these questions. Baltimorean's are in some way split into different groups by these questions. We explore a little in how they're split by these topics in this story.

### Groups of people with similar opinions

These are the percentages reported in the story. I won't repeat the writing there.

```{r}
srs_design %>%filter(presmosby>2,presmosby<88) %>%  group_by(interact(presmosby, samosby)) %>%
  summarise(prop = survey_prop()) %>% filter(samosby>2, samosby<88) %>% pull(prop) %>% sum

srs_design %>%filter(presmosby<3,presmosby<88) %>%  group_by(interact(presmosby, samosby)) %>%
  summarise(prop = survey_prop()) %>% filter(samosby>2, samosby<88) %>% pull(prop) %>% sum

srs_design %>%filter(presmosby>2,presmosby<88) %>%  group_by(interact(presmosby, scott)) %>%
  summarise(prop = survey_prop()) %>% filter(scott>2, scott<88) %>% pull(prop) %>% sum

srs_design %>%filter(presmosby>2,presmosby<88) %>%  group_by(interact(presmosby, policecom)) %>%
  summarise(prop = survey_prop()) %>% filter(policecom>2, policecom<88) %>% pull(prop) %>% sum

srs_design %>%filter(presmosby>2,presmosby<88) %>%  group_by(interact(presmosby, ceo)) %>%
  summarise(prop = survey_prop()) %>% filter(ceo>2, ceo<88) %>% pull(prop) %>% sum

```



```{r}
poll_dat_mos<-D_poll_results %>% filter(presmosby<5, samosby<5) %>% 
  mutate(samosby=case_when(samosby%in%c(1,2)~1,samosby%in%c(3,4)~2),
         presmosby=case_when(presmosby%in%c(1,2)~1,presmosby%in%c(3,4)~2),
         pres_mosby_j = jitter(presmosby, 1.2),
         samosby_j = jitter(samosby, 1.2),
         samosby=case_when(samosby==1~"dislikes_sa",samosby==2~"likes_sa"),
         presmosby=case_when(presmosby==1~"Dislikes Nick Mosby",presmosby==2~"Likes Nick Mosby"),
 
        ) %>% select(pres_mosby_j, samosby_j, presmosby, samosby) 

poll_dat_mos %>% 
  ggplot(aes(x = pres_mosby_j, y = samosby_j))+geom_point()+
  scale_x_discrete(limits=c("Disapproves of Nick Mosby","Approves of Nick Mosby"))+
  scale_y_discrete(limits=c("Disapproves of Marilyn Mosby","Approves of Marilyn Mosby"))+
  labs(x = "Nick Mosby", y = "Marilyn Mosby", title = "Do you approve of the job X is doing")

```


### What's wrong with the Baltimore Police

Like before, this is the calculation of the percentages that go along with the figures in the story.

These correspond to respondents attitudes on greater prosecution of nonviolent offenses and a variety of other crime-reduction measures polled.

```{r}
srs_design %>%filter(prosecution%in%c(3,4)) %>%  group_by(interact(prosecution, harsh)) %>%
  summarise(prop = survey_prop()) %>% filter(harsh%in%c(3,4)) %>% pull(prop) %>% sum

srs_design %>%filter(prosecution%in%c(1,2)) %>%  group_by(interact(prosecution, harsh)) %>%
  summarise(prop = survey_prop()) %>% filter(harsh%in%c(3,4)) %>% pull(prop) %>% sum

srs_design %>%filter(prosecution%in%c(3,4)) %>%  group_by(interact(prosecution, allocate)) %>%
  summarise(prop = survey_prop()) %>% filter(allocate%in%c(3,4)) %>% pull(prop) %>% sum

srs_design %>%filter(prosecution%in%c(1,2)) %>%  group_by(interact(prosecution, allocate)) %>%
  summarise(prop = survey_prop()) %>% filter(allocate%in%c(3,4)) %>% pull(prop) %>% sum

srs_design %>%filter(pol_fund%in%c(3)) %>%  group_by(interact(pol_fund, allocate)) %>%
  summarise(prop = survey_prop()) %>% filter(allocate%in%c(3,4)) %>% pull(prop) %>% sum
```

We also examine how Baltimore's view of the issues in the police department (are they systemic or limited to a few bad apples) depends on race, age, and ideology. 

All confidence intervals (read, significance testing) here is done at the 90% (10%) level. We're using a slightly lowered probability level because we'd like to highlight potentially interesting differences as different. 

```{r}
plot_data_1<-srs_design %>% group_by(interact(race_cat, ideo, age), few_dept) %>% 
  summarise(prop=survey_prop(vartype="ci", level = .9))#

plot_data_2_r<-plot_data_1%>%
  filter(race_cat%in%c(1,2), ideo %in%c(1,2,3), few_dept%in%c(1,2),age!=1 ) %>% na.omit %>% 
  mutate(race_cat = as_factor(race_cat), ideo = as_factor(ideo), few_dept = as_factor(few_dept),
         age = as_factor(age)) %>% group_by(race_cat, age) %>% 
  mutate(chart_group = cur_group_id()) %>% ungroup %>% 
  mutate(cat_title = str_c(race_cat, ", ",age),
         prop = prop*100,prop_low = prop_low*100, prop_upp=prop_upp*100) 

plot_data_2_flou<-plot_data_2_r%>% 
  pivot_wider(names_from = few_dept, values_from = c(prop, prop_low, prop_upp)) %>% repair_names()

plot_data_2_r %>% filter(race_cat == "White", age=="45-54", ideo%in%c("Conservative","Progressive"))
```

Here is the plot that corresponds to the department/race/age/ideology chart in the story. You can see the confidence intervals plotted on this chart that correspond to the significance annotations put online. 

```{r}
names(plot_data_2_flou)[6]<-"A few bad officers"
names(plot_data_2_flou)[7]<-"Departmentwide"
names(plot_data_2_flou)[8]<-"few_bad_l"
names(plot_data_2_flou)[9]<-"dept_wide_l"
names(plot_data_2_flou)[10]<-"few_bad_h"
names(plot_data_2_flou)[11]<-"dept_wide_h"

plot_data_2_r%>% 
  ggplot(aes(x = ideo, y = prop, fill = few_dept))+
  geom_bar(position="dodge", stat="identity")+
  geom_errorbar(aes(ymin=prop_low, ymax = prop_upp, x = ideo), position="dodge")+
  facet_wrap(~race_cat+age)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Heatmap

Finally, here we present the heatmap that goes with the story. I haven't seen anyone use a heatmap to present a survey before, but I think it's an interesting way to go about things. It's worth mentioning that this plot is really just an illustration of what's happening in the survey for a variety of reasons. 

First off, we're not taking weights into account here. Normally, that's a pretty big deal, however, because the sampled population happened to be quite close to Baltimore's population on the measures weighted for, the largest sample weight in the survey is 2.63. So, it's a pretty close representation.

We're also aggregating individual opinions to present opinions in blocks of 50 for the chart online. This also introduces a bit of smoothing into the chart. 

However, the lessons gleaned from the heatmap are similar to the lessons one sees upon a close examination of the data, but without having to examing 30-some variables (or charts) by hand. 

We also use a dendogram-based sorting algorithm to sort columns and rows in a way that visually groups similar blocks of respondents and questions to make reading the heatmap easier than it would be otherwise.

Before that, though, here are the calculations for the total dissatisfaction and Santelises statistics.
```{r}
D_for_diff=srs_design %>%mutate(
  opt = case_when(opt%in%c(1,2)~2,
                  opt%in%c(3,4)~1),
  
  scott = case_when(scott%in%c(1,2)~2,
                    scott%in%c(3,4)~1),
  
  presmosby = case_when(presmosby%in%c(1,2)~2,
                        presmosby%in%c(3,4)~1),
  
  samosby = case_when(samosby%in%c(1,2)~2,
                      samosby%in%c(3,4)~1),
  
  police = case_when(police%in%c(1,2)~2,
                     police%in%c(3,4)~1),
  
  policecom = case_when(policecom%in%c(1,2)~2,
                        policecom%in%c(3,4)~1),
  
  ceo = case_when(ceo%in%c(1,2)~2,
                  ceo%in%c(3,4)~1),
  
  scott_fav = case_when(scott_fav%in%c(1,2)~2,
                        scott_fav%in%c(3,4)~1),
  
  afford = case_when(afford%in%c(1,2)~1,
                     afford%in%c(3)~2),  
  
  litter = case_when(litter%in%c(1,2)~1,
                     litter%in%c(3)~2),  
  
  pub_transit = case_when(pub_transit%in%c(1,2)~1,
                          pub_transit%in%c(3)~2),  
  
  job_opp = case_when(job_opp%in%c(1,2)~1,
                      job_opp%in%c(3)~2),  
  
  crime = case_when(crime%in%c(1,2)~1,
                    crime%in%c(3)~2), 
  
  taxrate = case_when(taxrate%in%c(1,2)~1,
                      taxrate%in%c(3)~2), 
  
  grade = case_when(grade%in%c(1,2,3,4)~1,
                    grade%in%c(5)~2), 
  
  condition = case_when(condition%in%c(1,2)~1,
                        condition%in%c(3)~2), 
  
  unsafe = case_when(unsafe%in%c(1,2)~1,
                     unsafe%in%c(3)~2), 
  
  fund_edu = case_when(fund_edu%in%c(1,2)~1,
                       fund_edu%in%c(3)~2), 
  
  standards = case_when(standards%in%c(1,2)~1,
                        standards%in%c(3)~2), 
  
  local_sch = case_when(local_sch%in%c(1,2)~1,
                        local_sch%in%c(3)~2), 
  
  teacher = case_when(teacher%in%c(1,2)~1,
                      teacher%in%c(3)~2), 
  
  safe_city = case_when(safe_city%in%c(1,2)~2,
                        safe_city%in%c(3,4)~1), 
  
  safe_nb = case_when(safe_nb%in%c(1,2)~2,
                      safe_nb%in%c(3,4)~1), 
  
  misconduct = case_when(misconduct%in%c(1,2)~2,
                         misconduct%in%c(3,4)~1), 
  
  treat = case_when(treat%in%c(1,2)~2,
                    treat%in%c(3,4)~1), 
  
  protect = case_when(protect%in%c(1,2)~2,
                      protect%in%c(3,4)~1), 
  
  few_dept = case_when(few_dept%in%c(1)~2,
                       few_dept%in%c(2)~1), 
  
  allocate = case_when(allocate%in%c(1,2)~1,
                       allocate%in%c(3,4)~2), 
  
  harsh = case_when(harsh%in%c(1,2)~1,
                    harsh%in%c(3,4)~2), 
  
  prosecution = case_when(prosecution%in%c(1,2)~1,
                          prosecution%in%c(3,4)~2), 
  
  safe_streets = case_when(safe_streets%in%c(1,2)~1,
                           safe_streets%in%c(3,4)~2), 
  
  presence = case_when(presence%in%c(1,2)~1,
                       presence%in%c(3,4)~2), 
  gun = case_when(gun%in%c(1,2)~1,
                  gun%in%c(3,4)~2), 
  
  pol_fund = case_when(pol_fund%in%c(1,2)~1,
                       pol_fund%in%c(3)~2)
) 

D_new<-D_for_diff$variables%>% select(-`case#`, -weight, - phone_type, -agequal, -reside, - condition,
                               -unsafe, -vacant_other, -reason_other,-like_oe, -starts_with("reason"),
                               -starts_with("leave"), -starts_with("race"), -nr_zip_01, 
                               -nr_like_01, -qzipcode, -income,-income_6, - educ, - age,
                               -year, -year_age, -grade, -lived, -ends_with("_2"), -gender, -EDUC_3,
                               -regvote, -ideo, -raised, - parent, -track, -opt, - leaving,-vacant,
                               
                               ) %>% 
  mutate_all(as.numeric) %>% as.matrix

D_new[is.na(D_new)]<-0
D_new[D_new>10]<-0

row_dist <- as.dist(1-cor(t(D_new), method = "pearson"))
col_dist <- as.dist(1-cor(D_new, method = "pearson"))

col_hc <- hclust(col_dist, method = "complete")
row_hc <- hclust(row_dist, method = "complete")
```

```{r}
length(which(D_new==2))/length(D_new)
srs_design %>% filter(parent%in%c(1,2)) %>% group_by(parent, ceo) %>% summarise(survey_prop()) %>% filter(ceo==88)
```


```{r}
heatmap.2(D_new, Rowv=dendsort(as.dendrogram(row_hc), isRevers=TRUE, type="average"), 
          Colv=dendsort(as.dendrogram(col_hc), type="average"),trace = "none",
          col=brewer.pal(9, "Blues"),
          dendrogram = "none",
          key = FALSE,
          lwid=c(0.2,5),
          lhei=c(0.2,5))

```

