---
title: "Banner General Poll"
output: github_document
date: "2022-09-20"
---
## General Fall 2022 poll analysis

Another poll, another analysis. And another chance to learn something about the electorate in Maryland while bringing some new stats methods to the problem, as well. 

First we load in the data and the libraries that we'll be using for it.

```{r setup, include=FALSE}
library(tidyverse)
library(srvyr)
library(survey)
library(vegan)
library(haven)

#we filter the data into a separate likely voter file because we'll be calling on this so often.
D_poll_results = read_sav("~/Desktop/banner_projects/banner_poll_stories/banner_general_poll/md_statewide_dat.SAV")
srs_design<-D_poll_results %>% as_survey_design(ids =1, weight = weight)
srs_design_lv <- srs_design %>% filter(LV_FINAL_SCREEN==1)
```


## Abortion

I touch on this in the post a bit, but this poll analysis is really a companion piece to the primary poll analysis I did a few months ago, in much the same way that a general election is a companion to a midterm election. A lot of the topics motivating voters are the same, and that shows up in the interesting data questions. 

We start by looking at the views Marylanders hold about abortions, and whether they've been differentially motivated by Dobbs v. Jackson ruling. We find that they have been, inways that I go into in the piece. Below is the data we write out for the Flourish map, as well as the single statistics that we call on.

```{r}
srs_design_lv %>% filter(ideo%in%c(1,2,3) ,sc_motivate%in%c(2,3), gender%in%c(1,2)) %>% group_by(interact(gender,ideo),sc_motivate) %>%
  summarise(survey_prop("ci", .9)) %>% 
  mutate(sc_motivate = as_factor(sc_motivate),ideo = as_factor(ideo), gender = as_factor(gender)) %>% 
  ggplot(aes(x = ideo, fill = sc_motivate, y = coef))+geom_bar(stat="identity", position = "dodge")+
  geom_errorbar(aes(ymin= `_low`, ymax = `_upp`,x = ideo), position="dodge")+facet_grid(~gender)

```

```{r}
srs_design_lv  %>% group_by(ideo, gender, sc_motivate) %>% summarise(survey_prop()) %>% ungroup%>%mutate_all(as_factor) %>% filter(ideo == "Progressive", gender =="Female", sc_motivate=="More motivated")
```

```{r}
srs_design_lv  %>% group_by(ideo, gender, sc_motivate) %>% summarise(survey_prop()) %>% ungroup%>%mutate_all(as_factor) %>% filter(ideo == "Conservative", gender =="Female", sc_motivate%in%c("Made no difference", "Less motivated")) %>% pull(coef) %>% as.character %>% as.numeric %>% sum
```

```{r}
srs_design_lv  %>% group_by(ideo, gender, sc_motivate) %>% summarise(survey_prop()) %>% ungroup%>%mutate_all(as_factor) %>% filter(ideo == "Conservative", gender =="Male", sc_motivate%in%c("Made no difference", "Less motivated")) %>% pull(coef) %>% as.character %>% as.numeric %>% sum
```

```{r}
srs_design_lv  %>% group_by(ideo, gender, sc_motivate) %>% summarise(survey_prop()) %>% ungroup%>%mutate_all(as_factor) %>% filter(ideo == "Moderate", gender =="Male", sc_motivate=="More motivated") 
```

```{r}
srs_design_lv  %>% group_by(ideo, gender, sc_motivate) %>% summarise(survey_prop()) %>% ungroup%>%mutate_all(as_factor) %>% filter(ideo == "Moderate", gender =="Female", sc_motivate=="More motivated") 
```

## Race relations

There was an interesting high-level result that white respondents had worse feelings about race relations than Black respondents, which, if true, is an interesting point that supports particular ideological views. However, if you look a little deeper into it, it doesn't seem to be true in the way it appears at the high level. 

Below, I show the simple hypothesis test we perform in the article to test for the high-level difference, as well as the closer examination that we talk about in the article. I also calculate what percentage of the Black electorate falls into the categories that seem to show a difference. 

```{r}
srs_design %>%filter(race%in%c(1,2)) %>%  group_by(race,race_rel_2) %>% summarise(survey_prop()) %>% mutate_all(as_factor) %>% 
  filter(race_rel_2=="Good")
```

```{r}
srv_count_by_race_for_test<-srs_design %>%filter(race%in%c(1,2)) %>%  group_by(race,race_rel_2) %>% summarise(survey_count(.))

n_white_good<-srv_count_by_race_for_test %>% mutate(race = as_factor(race), race_rel_2 = as_factor(race_rel_2)) %>% filter(race=="White", race_rel_2=="Good") %>% unique() %>% pull(n)

n_black_good<-srv_count_by_race_for_test %>% mutate(race = as_factor(race), race_rel_2 = as_factor(race_rel_2)) %>% filter(race=="Black", race_rel_2=="Good") %>% unique() %>% pull(n)

n_tot_black <- srs_design %>% group_by(race) %>% summarise(survey_count(.)) %>% mutate(race = as_factor(race)) %>% filter(race=="Black") %>% pull(n) %>% unique

n_tot_white <- srs_design %>% group_by(race) %>% summarise(survey_count(.)) %>% mutate(race = as_factor(race)) %>% filter(race=="White") %>% pull(n) %>% unique

test<-prop.test(c(n_white_good, n_black_good), c(n_tot_white,n_tot_black)) # the difference is significant but small
test
```
```{r}
srs_design %>%filter(race%in%c(1,2)) %>%  group_by(interact(race,ideo),race_rel_2) %>% summarise(survey_prop()) %>%
  mutate(ideo = as_factor(ideo), race_rel_2 = as_factor(race_rel_2)) %>% filter(ideo=="Conservative",race_rel_2 =="Good")
```

```{r}
srs_design %>%filter(race%in%c(1,2)) %>%  group_by(interact(race,ideo),race_rel_2) %>% summarise(survey_prop("ci",.9))  %>%
  mutate(race = as_factor(race), race_rel = as_factor(race_rel_2), ideo = as_factor(ideo)) %>% select(-race_rel_2) %>% ungroup %>% 
  select(-`_low`,-`_upp`) %>% 
  pivot_wider(names_from = race, values_from = coef) %>% filter(race_rel=="Good") %>% 
  mutate(White = White*100, Black =Black*100) %>% 
  write_csv("~/Desktop/banner_projects/banner_poll_stories/banner_general_poll/race_rel_by_ideo.csv")
```

```{r}
srs_design %>%filter(race%in%c(1,2)) %>%  group_by(interact(race,ideo),race_rel_2) %>% summarise(survey_prop()) %>%
  mutate(ideo = as_factor(ideo), race_rel_2 = as_factor(race_rel_2)) %>% filter(ideo=="Conservative",race_rel_2 =="Bad")
```

```{r}
srs_design %>%filter(race%in%c(1,2)) %>%  group_by(interact(race,ideo),race_rel_2) %>% summarise(survey_prop()) %>%
  mutate(ideo = as_factor(ideo), race_rel_2 = as_factor(race_rel_2), race = as_factor(race)) %>% filter(ideo=="Conservative",race_rel_2 =="Good")
```

```{r}
srs_design %>%mutate(region_code=as_factor(region_code)) %>% filter(race%in%c(1,2), region_code=="Baltimore County") %>%  group_by(interact(race,region_code),race_rel_2) %>% summarise(survey_prop()) %>% mutate(race = as_factor(race), race_rel_2 = as_factor(race_rel_2)) %>%  filter(race=="Black", race_rel_2=="Good")


srs_design %>%filter(race%in%c(1,2)) %>%  group_by(interact(race,region_code),race_rel_2) %>% summarise(survey_prop("ci",.9))  %>%
  mutate(race = as_factor(race), race_rel = as_factor(race_rel_2), region_code = as_factor(region_code)) %>%
  select(-race_rel_2, -`_low`, -`_upp`) %>%pivot_wider(names_from= race, values_from = coef) %>% filter(race_rel=="Good") %>% 
  mutate(White = White*100, Black =Black*100) %>% 
  write_csv("~/Desktop/banner_projects/banner_poll_stories/banner_general_poll/race_rel.csv")
```
```{r}
srs_design %>%mutate(region_code=as_factor(region_code)) %>% filter(race%in%c(1,2), region_code=="Baltimore County") %>%  group_by(interact(race,region_code),race_rel_2) %>% summarise(survey_prop()) %>% mutate(race = as_factor(race), race_rel_2 = as_factor(race_rel_2)) %>%  filter(race=="White", race_rel_2=="Good")
```
```{r}
tot_black<-srs_design %>% mutate(race = as_factor(race)) %>% filter(race=="Black") %>% pull(weight) %>% sum

black_conservatives<-srs_design %>% mutate(race = as_factor(race), ideo = as_factor(ideo)) %>% filter(race=="Black", ideo=="Conservative") %>% pull(weight) %>% sum

black_not_cons_balt_cou<-srs_design %>% mutate(race = as_factor(race), ideo = as_factor(ideo), region_code = as_factor(region_code)) %>% filter(race=="Black", ideo!="Conservative", region_code =="Baltimore County") %>% pull(weight) %>% sum

#30 %
(black_not_cons_balt_cou+black_conservatives)/tot_black
```

## Isomap

Lastly, we have what I find, as a methods-loving statistician, the most fun part of the analysis--some dimensionality reduction tricks. I mention this in the article a little bit, but it's worthwhile to have some kind of way to visualize a high-dimensional space as a lower dimensional space. Of course, a poll, in some ways, represents an ideal high-dimensional space with a lower-dimensional subspace representation. While there are many questions on the poll, each of which could be viewed as a dimension, there are far fewer "kinds of opinions." Many of these dimensions correlate with one another, but, intuitively, political opinions admit a representation.

In general, I'm partial to non-linear dimensionality reduction techniques. Compare how PCA does on the famous Swiss roll set to non-linear embeddings and you'll see what I mean. To go into probably a little too much detail, many dimensionality reduction techniques function in effectively the same way. You define a metric, calculate distances using that metric, do an eigenvalue decomposition, and reduce the dimension by picking a certain number of top eigenvalues. 

Isomap uses a path-based distance. You define a neighborhood around each point, connect it to its neighbors, and the calculate distances between points by counting the length of the path from one point to another. The goal with that is to preserve the local structure of the data.

We apply this method to the polling data to interesting effect. For one, it's clear that there is local structure that the data is pulling out. But we also learn some interesting things from viewing the data in this way that I talk about in the article.

```{r}
D_for_diff_lv<-srs_design_lv$variables %>%mutate(
  hoganjob = case_when(hoganjob%in%c(1,2)~2,
                        hoganjob%in%c(3,4)~1),
  
  mdecon = case_when(mdecon%in%c(1)~1,
                     mdecon%in%c(2)~2),
  
  price = case_when(price%in%c(1)~2,
                    price%in%c(2,3)~1),
  
  bidenjob = case_when(bidenjob%in%c(3,4)~2,
                       bidenjob%in%c(1,2)~1),
  
  trumpfav = case_when(trumpfav%in%c(1,2)~1,
                       trumpfav%in%c(3,4)~2),
  
  racism = case_when(racism%in%c(1,2)~1,
                     racism%in%c(3,4)~2),
  
  crt = case_when(crt%in%c(1,2)~1,
                  crt%in%c(3,4)~2),
  
  vaccine = case_when(vaccine%in%c(1,2)~1,
                      vaccine%in%c(3,4)~2),
  
  parents = case_when(parents%in%c(1,2)~1,
                      parents%in%c(3,4)~2),  
  
  # crime = case_when(crime%in%c(1,2)~1,
  #                   crime%in%c(3)~2),  excluding crime from this analysis because i don't think there's a way to justify it with what we have
  
  race_rel = case_when(race_rel%in%c(1,2)~1,
                       race_rel%in%c(3,4)~2),  
  
  covid_over = case_when(covid_over%in%c(1,2)~2,
                         covid_over%in%c(3)~1),
  
  pubsafety = case_when(pubsafety%in%c(0)~2,
                         pubsafety%in%c(1)~1),
  
  econtax = case_when(econtax%in%c(0)~2,
                      econtax%in%c(1)~1),
  
  pubsch = case_when(pubsch%in%c(0)~2,
                     pubsch%in%c(1)~1),

  infra = case_when(infra%in%c(0)~2,
                    infra%in%c(1)~1),
  
  pubtransit = case_when(pubtransit%in%c(0)~2,
                         pubtransit%in%c(1)~1),
  
  environment = case_when(environment%in%c(0)~2,
                          environment%in%c(1)~1),
  
  housing = case_when(housing%in%c(0)~2,
                      housing%in%c(1)~1),
  
  issue_oth = case_when(issue_oth%in%c(0)~2,
                        issue_oth%in%c(1)~1),
  
  sc_motivate = case_when(sc_motivate%in%c(1,2)~2,
                          sc_motivate%in%c(3)~1),
  
  moore_fav = case_when(moore_fav%in%c(3,4)~2,
                        moore_fav%in%c(1,2)~1),
  
  cox_fav = case_when(cox_fav%in%c(3,4)~2,
                      cox_fav%in%c(1,2)~1),
  
  interest = case_when(interest%in%c(1,2)~1,
                       interest%in%c(3,4)~2),
  
  #not including ideology or favoriate candidate questions because we can't really code them in the way that i do. 
  
  marijuana = case_when(marijuana%in%c(2)~2,
                        marijuana%in%c(1)~1),
  
  expunge = case_when(expunge%in%c(2)~2,
                        expunge%in%c(1)~1)
) %>% as_survey_design(ids =1, weight = weight) %>% 
  select(hoganjob,mdecon,
    price,bidenjob, trumpfav,racism,crt,vaccine,parents,race_rel,covid_over,
    pubsafety,econtax,pubsch,infra, pubtransit,environment,housing,issue_oth,sc_motivate,
    moore_fav, cox_fav,marijuana, expunge, regvote,interest, ideo) 

d_for_dists<-D_for_diff_lv$variables %>% mutate_all(function(x)return(replace_na(x, 0)))

dataTable <- d_for_dists %>%select(-ideo, -regvote) %>%  as.matrix
dists_lv<-dist(dataTable)
D_iso<-isomap(dists_lv, k =8)
plot(D_iso$points[,c(1,2)], col =d_for_dists$regvote, pch = 16, cex = .6)
```


```{r}
srs_design_lv %>% group_by(interact(housing, environment, pubtransit, infra, pubsch, econtax, pubsafety)) %>% 
    summarise(survey_prop()) %>% mutate(last_three = pubsch +econtax+pubsafety) %>% filter(last_three>1, last_three<4) %>% pull(coef) %>% sum

srs_design_lv %>% group_by(interact(housing, environment, pubtransit, infra, pubsch, econtax, pubsafety)) %>% 
    summarise(survey_prop()) %>% mutate(last_three = pubsch +econtax+pubsafety) %>% filter(last_three==3) %>% pull(coef) %>% sum
```
